---
title: "Acticity Prediction"
author: "Marco Venturi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### Abstract
The dataset has the pourpose to predict which type of excercise a person is doing, the types of excercise is decide over the feature *classe*. The data were downloaded from a web source and cleaneed. The dataset is submitted by using a random forest model. 

### Downloading Data

```{r cars}
URL_training      <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
destfile_training <- "pml-training.csv"

URL_test      <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
destfile_test <- "pml-test.csv"

if(!file.exists(destfile_training)){
  download.file(URL_training, destfile_training)
}


if(!file.exists(destfile_test)){
  download.file(URL_test, destfile_test)
}
```

### Loading and Cleaning Data

Cleaning data

```{r pressure, echo=FALSE}
pml_training <-  read.csv("pml-training.csv")
pml_training <- pml_training[, c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140, 151:160)]

pml_testing <-  read.csv("pml-test.csv")
validating  <- pml_testing[, c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140, 151:160)]

```


### Slicing Data

Slicing data in train, test and validation set

```{r}
library(caret)
library(kernlab)

inTrain <- createDataPartition(y = pml_training$classe, p = 0.75, list = FALSE)

training <- pml_training[inTrain, ]
testing <- pml_training[-inTrain, ]
dim(training)
```
### Creating Model

Crating Random Forest model predictor.

```{r}
control          <- trainControl(method = "cv",number = 4, allowParallel = TRUE)
model            <- train(classe ~ ., data=training, method = "rf" , trControl = control)
predict          <- predict(model, testing)
confusionMatrix  <- confusionMatrix(factor(testing$classe), predict)

confusionMatrix
```

Printing errooe rate.

```{r}
accuracy <- confusionMatrix$overall[1]

error <- (1 - sum(predict == testing$classe) / length(predict))
error
```

### Making Predictions

Making prediction over validation set.

```{r}
result  <- predict(model, validating)
result
```
