---
title: "Course Project 1"
author: "Marco Venturi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Course Project week 1##  

###Looking in to data###  

This data set contain  information about about personal movement using activity monitoring devices such as a Fitbit,   Nike Fuelband, or Jawbone Up.    

**loading data**  

```{r}
library(ggplot2)

data <- read.csv("./data/activity.csv")

summary(data)

```

*Steps* has alot of NA values, about 2304, while *date* is not in date format.  Let's change *date* format and leave   NA values for now   

**data processing**  
```{r}
data$date <- as.Date(data$date, "%Y-%m-%d")

head(data,5)
```


###What is mean total number of steps taken per day###  

First it need to be calculate  the total number of steps per day(*step_day*).  


**steps per day**  
```{r}
step_day <- aggregate(steps ~ weekdays(date) , data, sum)

head(step_day, 5)

```



let's plot the histogram about *step_day*  

**histogram**  
```{r histogram, fig.height = 4, fig.width=7}
ggplot(step_day, aes(x =`weekdays(date)`, y = steps)) +
  geom_bar(stat = "identity",aes( fill = factor(`weekdays(date)`))) +
  ggtitle("Avg Steps Per Time Interval") +
  labs(x="Time interval", y="Avg Steps Per Day") +
  guides(fill = "none")
```
how much are *mean* and *median* of the total number of *steps* taken per day?  

**mean and median**  

```{r}
print(mean(step_day$steps))   #mean
print(median(step_day$step))  #median
```

###What is the average daily activity pattern?###  

**average number of steps taken all days across interval**
```{r}
step_5_min <- data
step_5_min$date <- weekdays(data$date)
head(step_5_min)
step_5_min <- aggregate(steps ~ date + interval , step_5_min, mean)
head(step_5_min)

```


**Line graph: average number of steps taken all days**
```{r linegraph, fig.height=4, fig.width=8}
ggplot(step_5_min, aes(x=interval, y = steps, color= date)) +
  geom_line() + 
  ggtitle("Avg Steps Per Time Interval") + 
  labs(x="Time interval one mesaurament each 5 minuts", y="average steps")
```

**5-minute interval on average across all the days that contains the maximum number of steps?**  
```{r}
max(step_5_min$step)
```

###Imputing missing values###  

**Number of missing values**
```{r}
na <- is.na(data)
nrow(data[na,])

```

**function for fill missing value in steps**

```{r}
fillStepBydayMean <- function(row, step_day) {
  
  row[,1] <-  step_day[step_day$day == row[,2], 1 ]
  
  return(row)
  
}

```

**filling missing data**
```{r,results='hide'}

colnames(step_day) <- c("day", "steps")

step_day$day <- factor(step_day$day)

fill_data <- data
fill_data$date <- factor(weekdays(as.Date(fill_data$date, "%Y-%m-%d")))

na_steps <- which(is.na(fill_data$steps))

i<-1L


for(index in na_steps) {
  fill_data[index, ] <- fillStepBydayMean(fill_data[index,], step_day )
}

```

```{r}
summary(fill_data)
head(fill_data)
```

```{r histogram_2 , fig.height = 4, fig.width=7}
step_day_filled_na <- aggregate(steps ~ date , fill_data, sum)


ggplot(step_day_filled_na, aes(x =date, y = steps)) +
  geom_bar(stat = "identity",aes( fill = date)) +
  ggtitle("Avg Steps Per Time Interval") +
  labs(x="Time interval", y="Avg Steps Per Day") +
  guides(fill = "none")

```

**mean and median fill_data**
```{r}
print(mean(step_day_filled_na$steps))   #mean
print(median(step_day_filled_na$step))  #median

```


###Are there differences in activity patterns between weekdays and weekends? ###


**Create a new factor variable in the data set with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.**


```{r}
fill_data$date <- weekdays(as.Date(data$date, "%Y-%m-%d"))
fill_data$weektype <- sapply(fill_data$date, function(x){
  if(x == "domenica" | x=="sabato" | x == "sunday"| x == "saturday"){
    result <- "weekend"
  }else{
    result <- "weekday"
  }
  
  return(result)
})

head(fill_data,5)
```


```{r}

step_5_min_filled <- fill_data
step_5_min_filled <- aggregate(steps ~ date + interval + weektype , step_5_min_filled, mean)
head(step_5_min_filled)

ggplot(step_5_min_filled, aes(x=interval, y = steps, color= date)) +
  geom_line() +
  facet_grid(.~ weektype) +
  ggtitle("Avg Steps Per Time Interval") + 
  labs(x="Time interval one mesaurament each 5 minuts", y="average steps")

```