---
title: "Storm Cosequence"
author: "Marco Venturi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis
  
This report study the differencies between all US countries about how much it cost different category of  storm events and how much fatalities those events cause.  
The questions that i tried to answer are :  

1. Across the United States, which types of events are most harmful with respect to population health?  
  
2.Across the United States, which types of events have the greatest economic consequences?  
  
In the firt answer i decide to analize *FATALITIES* and *EVTTYPE*, because i thought that the worst case is the case with more *FATALITIES*instead of *INJURIES*.  

In the second answer i decide to analize *CROPDMG* and *PROPDMG* for understand how much it cost differents types over different costs.

## Reading Data  
  
The zip folder was unziped using R.utils library, in specific *bunzip2()* function, where the output csv file is called as is folder, *repdata_data_StormData.csv* , then the new file was read with *read.csv* function.

```{r}

#New csv file is create inside data folder
library(R.utils)
library(ggplot2)

bunzip2("./data/repdata_data_StormData.csv.bz2", "repdata_data_StormData.csv", remove = FALSE, skip = TRUE)
df <- read.csv("./data/repdata_data_StormData.csv")

```


## Data Processing  

### 1) which types of events are most harmful with respect to population health?  

The first question was answered by sum all *FATALITIES* divide by *EVTTYPE* and adding a column to this new data set for checking if there where any Unknow  events, this was possible by checking unique values of *EVTTYPE*, because if the event is not Know the *EVTTYPE* is describe as *"^SummarY *" (regex)**.

```{r}
sum_fatalites_over_evtype <- aggregate(FATALITIES ~ EVTYPE, df, FUN = sum)

sum_fatalites_over_evtype$know_event <- sapply(sum_fatalites_over_evtype$EVTYPE, function(x){
  
  if(grepl("^Summary*", x)){
    result <- "Unknow Event"
  }else{
    result <- "Know Event"
  }
  return(result)
  
})

```

In the end i decide to order the data set for plot just the first 10th event over 985 types  

```{r}
sum_fatalites_over_evtype <- sum_fatalites_over_evtype[order(sum_fatalites_over_evtype$FATALITIES , decreasing = TRUE),]

head(sum_fatalites_over_evtype ,10)

```


### 2) which types of events have the greatest economic consequences?  
  
The second question was answered by sum all *CROPDMG* and *PROPDMG* divide by *EVTTYPE* and adding a column to this new data set for checking if there where any Unknow  events, this was possible by checking unique values of *EVTTYPE*, because if the event is not Know the *EVTTYPE* is describe as *"^SummarY *" (regex)**.

```{r}
sum_damage <- aggregate(CROPDMG + PROPDMG ~ EVTYPE, df, sum )

head(sum_damage)

sum_damage$know_event <- sapply(sum_damage$EVTYPE, function(x){
  
  if(grepl("^Summary*", x)){
    result <- "Unknow Event"
  }else{
    result <- "Know Event"
  }
  return(result)
  
})

```

In the end i decide to order the data set for plot just the first 10th event over 985 types  

```{r}
sum_damage <- sum_damage[order(sum_damage$`CROPDMG + PROPDMG` , decreasing = TRUE),]

head(sum_damage, 10)

```

## Results

### 1)Across the United States, which types of events (as indicated in the \color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?  
  
Tornado seems to be the most dangearous event across all country, while other event under the 5TH are less dangeorus with around hundreds of deads

```{r}
ggplot(sum_fatalites_over_evtype[1:10, ], aes(y = FATALITIES, x = EVTYPE )) +
  geom_bar(stat = "identity",aes( fill = factor(EVTYPE))) +
  ggtitle("First 10th event type across fatalities") +
  labs(x="Event type", y="Fatalities") +
  guides(fill = "none")

```

### 2)Across the United States, which types of events (as indicated in the \color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?  
  
  
Tornado seems to be the most expensive event across all country, while other event are still expesive but less then 1.500.000$

```{r}
ggplot(sum_damage[1:10, ], aes(y = `CROPDMG + PROPDMG`/10^6, x = EVTYPE )) +
  geom_bar(stat = "identity",aes( fill = factor(EVTYPE))) +
  ggtitle("First 10th event type across damage divide by 10^6") +
  labs(x="Event type", y="damage") +
  guides(fill = "none")

```
